Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-04-22 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
update

%%%%%%%%%%%% 
\subsection{Initial Model Validation}\label{section:results_initial_model_validation}
Numerical simulations \citep[based on lab experiments due to][]{evans_1994,dawe_2008} were conducted to demonstrate the model capability to capture viscous crossflow during immiscible displacement in heterogeneous porous media. The 2D domain, shown in Fig.~\ref{fem_cv_represent_a}, is $4$ $\times$ $2$ unit-length and fully saturated with fluid 2 (VR=1). Fluid 1 is injected from the left-hand side of the domain with constant velocity of $u=1$. \green{No normalflux boundary conditions are implied across the upper and lower borders}. The domain consists of four regions in which each quarter is represented by a permeability value, Fig.~\ref{fem_cv_represent_a}(a). The spatial permeability distribution creates a rough pressure field, \ie pressure gradient is discontinuous across the interface between different permeability zones, thus the fluid flows to the larger permeability region. %Figure \ref{fem_cv_represent_a}(b-c)describe the saturation field at different times.

@@

\medskip
Here, two numerical simulations were conducted with fixed and adaptive mesh resolutions to qualitatively assess model capability to capture fingers dynamics. In order to trigger the formation of fingers, regions with sharp permeability gradient were introduced. The computational domain, Fig.~\ref{fig:testcase_heter_domain}a, consists of a rectangular geometry fully saturated with a fluid (except by a squared region containing 50$\%$ of a second fluid -- wetting fluid phase, Fig.~\ref{fig:testcase_heter_domain}b) and divided into 5 regions with prescribed \green{synthetic permeability distribution with values between $1$ and $5$}. A no-flux boundary condition was imposed across upper and lower borders, whilst pure wetting fluid phase $\left(\text{\ie} S_{1}=1\right)$ is driven into the domain from the left-hand side face at velocity $u_{1}=1$. Viscosity ratio was set to 10.

>>>>

%%%%%%%%%%%% 
\subsection{Initial Model Validation}\label{section:results_initial_model_validation}
Numerical simulations \citep[based on lab experiments due to][]{evans_1994,dawe_2008} were conducted to demonstrate the model capability to capture viscous crossflow during immiscible displacement in heterogeneous porous media. The 2D domain, shown in Fig.~\ref{fem_cv_represent_a}, is $4$ $\times$ $2$ unit-length and fully saturated with fluid 2 (VR=1). Fluid 1 is injected from the left-hand side of the domain with constant velocity of $u=1$. \red{No normal flux boundary conditions are implied across the upper and lower borders}. The domain consists of four regions in which each quarter is represented by a permeability value, Fig.~\ref{fem_cv_represent_a}(a). The spatial permeability distribution creates a rough pressure field, \ie pressure gradient is discontinuous across the interface between different permeability zones, thus the fluid flows to the larger permeability region. %Figure \ref{fem_cv_represent_a}(b-c)describe the saturation field at different times.

@@

\medskip
Here, two numerical simulations were conducted with fixed and adaptive mesh resolutions to qualitatively assess model capability to capture fingers dynamics. In order to trigger the formation of fingers, regions with sharp permeability gradient were introduced. The computational domain, Fig.~\ref{fig:testcase_heter_domain}a, consists of a rectangular geometry fully saturated with a fluid (except by a squared region containing 50$\%$ of a second fluid -- wetting fluid phase, Fig.~\ref{fig:testcase_heter_domain}b) and divided into 5 regions with prescribed \red{synthetic permeability distribution with values between $1$ and $5$}. A no-flux boundary condition was imposed across upper and lower borders, whilst pure wetting fluid phase $\left(\text{\ie} S_{1}=1\right)$ is driven into the domain from the left-hand side face at velocity $u_{1}=1$. Viscosity ratio was set to 10.
